---
title: "HW1"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
- What to hand in: soft copies of a notebook (Rmd file) with your answers (you can use this one as a template), and the corresponding knitted pdf. 
- When: Due at noon, Thursday September 6, 2018.
- How: Information on how to submit the files will be given on moodle. 

We will use data from the gapminder package, let's load it. 
```{r setup}
library(tidyverse)
install.packages("gapminder")
library(gapminder)
```

A look at the data:
```{r}
gapminder
```

A simple base R plot of life expectancy against log(GDP):
```{r}
# Base R plot
plot(lifeExp ~ log(gdpPercap), data = gapminder)
```



# HW exercises: 
use ggplot to make the graphs requested in (1) and (2). Throughout, avoid repetition of code.
Goal: gain experience with ggplot and reading help files/using google searches to get help on (arguments of) R functions. 

1. make one scatter plot using ggplot with
- Life expectancy on the y-axis, label axis as ``Life expectancy''
- GDP per capita on the x-axis, use a log-transform of the x-axis while still displaying the (unlogged) GDP values 
- shapes to indicate the continent, add a legend
- color to indicate the year, do NOT add a legend for that

2. add two geom_smooth layers (a) and (b) to the plot you made in (1) where
(a) is fitted to the entire dataset and uses default settings for the smoother used and plotting settings
(b) is fitted to data in Africa only, provides a 99% confidence interval, is plotted in red (both point estimates as well as confidence interval), and is transparent (so if it overlaps with (a), (a) is still visible). 
Hint: to fit to data in Africa only, consider using "data = filter(gapminder, continent == "Africa")"

```{r}
# 1

?plot.ggplot

# y-axis = Life expectancy
# x-axis = GDP per capita (pay attention to log vs. unlogged)
# shapes to indicate the continent, add a legend
# color to indicate year, no legend 

ggplot(data=gapminder) +
  geom_point(mapping = aes (x=log(gdpPercap), y=lifeExp, shape=continent, color=year)) +
  guides(color=F) +
  xlab ("GDP per capita") +
  ylab ("LifeExpectancy")

```
```{r}
#2

# add two geom_smooth layers (a) and (b) to the plot you made in (1) where
# (a) is fitted to the entire dataset and uses default settings for the smoother used and plotting settings
# (b) is fitted to data in Africa only...
# provides a 99% confidence interval...
# is plotted in red (both point estimates as well as confidence interval),...
# and is transparent (so if it overlaps with (a), (a) is still visible). ...
# Hint: to fit to data in Africa only, consider using "data = filter(gapminder, continent == "Africa")"

ggplot(gapminder) +
  geom_point(mapping = aes (x=log(gdpPercap), y=lifeExp, shape=continent, color="red")) +
  geom_smooth(aes(log(gdpPercap), lifeExp)) +
  geom_smooth(data = filter(gapminder, continent == "Africa"), aes(log(gdpPercap), lifeExp), level=0.99, color="red", fill="red", method=lm) +
  xlab ("GDP per capita") +
  ylab ("Life Expectancy") +
  guides(color=F)
  
```


